CUBE: A Memory-lite Cosmological $N$-body Algorithm

### 2017-09-29 ms_caf/ms_caf.tex
corrected numbers in table 1 -- fixed equivalence bugs.

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
 \caption{particles.}
\label{fig.particles}
\end{figure}

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
 \caption{Displacement.}
\label{fig.dsp}
\end{figure}

@@

\subsubsection{Coarse mesh arrays}
On coarse mesh, {\tt rho\_c}, {\tt vfield} and force kernel {\tt kern\_c} should always be kept. They are usually configured to be 4 times coarser than fine grids and particle number density. In this case, each coarse grid, 3D {\tt vfield} takes only 12 bytes, and 64 particles will at least take 384 bytes ($n_\chi=n_\nu=1$ case). Similarly, pencil-FFT arrays, coarse force arrays etc. are memory-light and temporary. The total coarse mesh arrays take less than 10\% of the memory of {\tt xvp}.

\subsubsection{Fine mesh arrays}
On the local fine mesh, only a force kernel array needs to be kept. The fine mesh density arrays, force field arrays are temporary. Since {\tt xvp\_new}, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements. The total fine mesh arrays take less than 5\% of the memory of {\tt xvp}.

@@

By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=4$, ${\tt ncb}=6$, $\epsilon_{\rm image}=10\%$, $\epsilon_{\rm tile}=100\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 38.3 GB of memory, corresponding to 10.6 bpp. This can be done on most of the supercomputers, even laptops.

\begin{table}[]
\centering
\caption{Memory layout for a certain configuration}
\label{t.memory}
\begin{tabular}{llrrr}
\hline
& & \multicolumn{3}{c}{Memory usage}\\
\cline{3-5}
Type     & Array & /GB & /bpp & Percentage \\
\hline
Particles  & \tt xvp          & 29.9   & 8.24   & 87.4\%     \\
           & \tt xvp\_new     & \underline{1.69}   & \underline{0.465}  & \underline{4.93\%}`     \\
           & \tt PID          & 0      & 0      &    0\%     \\
           & Subtotal         & 31.5   & 8.71   & 92.4\%     \\
\hline
Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.870\%    \\
           & \tt vfield       & 0.889  & 0.245  & 2.60\%     \\
           & \tt kern\_c      & 0.340  & 0.0939 & 0.996\%    \\
           & \tt force\_c     & (0.690)  & (0.190)  & (2.02\%)    \\
           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0411\% \\
           & Pencil-FFT       & (0.454)  & (0.125) & (1.33\%)      \\
           & Subtotal         & 1.55   & 0.429 & 4.55\%     \\
\hline
Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 3.09\%      \\
           & \tt force\_f     & (1.63) & (0.450) & (4.77\%)     \\
           & Fine-FFT     &   (1.41) & (0.389) & (4.13\%)     \\
           & Subtotal      & 1.06   & 0.292 & 3.09\%      \\
\hline
Total &                      & 34.2 & 9.43 & 100\%\\
\hline
\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 63.6\%\\
\hline
\end{tabular}
\end{table}

@@

Results of {\tt izipx=1}

Results of {\tt izipx=2}

\begin{figure}
\centering
  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
 \caption{power spectrum.}
\label{fig.power}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1.1\linewidth]{diff_power.pdf}
 \caption{power spectrum and cross correlation.}
\label{fig.ccc}
\end{figure}

>>>>

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{overplotparticles.pdf}
 \caption{Difference between using {\tt CUBEP3M} (grey dots) and {\tt CUBE}-x1v1 (red dots). The box shows a projected $15.625^2\times 3.90625\ ({\rm Mpc}/h)^3$ region of the simulation. }
\label{fig.particles}
\end{figure}

\begin{figure}[t]
\centering
  \includegraphics[width=0.95\linewidth]{ddsp_pcdf_256_box80.pdf}
 \caption{Error of the displacement of particles.}
\label{fig.dsp}
\end{figure}

@@

\subsubsection{Coarse mesh arrays}
On coarse mesh, {\tt rho\_c}, {\tt vfield} and force kernel {\tt kern\_c} should always be kept. They are usually configured to be 4 times coarser than fine grids and particle number density. In this case, each coarse grid, 3D {\tt vfield} takes only 12 bytes, and 64 particles will at least take 384 bytes ($n_\chi=n_\nu=1$ case). Similarly, pencil-FFT arrays, coarse force arrays etc. are memory-light and temporary.

\subsubsection{Fine mesh arrays}
On the local fine mesh, only a force kernel array needs to be kept. The fine mesh density arrays, force field arrays are temporary. Since {\tt xvp\_new}, coarse force arrays, pencil-FFT arrays are also temporary and are not used in any calculation simultaneously, they can be overlapped in memory by using equivalent statements.

@@

By using {\tt CUBE}, with parameters $n_\chi=n_\nu=1$, ${\tt nc}=384$, ${\tt nnt}=3$, ${\tt ncb}=6$, $\epsilon_{\rm image}=5\%$, $\epsilon_{\rm tile}=200\%$ and $\left\langle {\tt nplocal} \right\rangle=1536^3$, we use about 35.6 GB of memory, corresponding to 9.84 bpp. This can be done on most of the supercomputers, even modern laptops.

\begin{table}[]
\centering
\caption{Memory layout for a certain configuration}
\label{t.memory}
\begin{tabular}{llrrr}
\hline
& & \multicolumn{3}{c}{Memory usage}\\
\cline{3-5}
Type     & Array & /GB & /bpp & Percentage \\
\hline
Particles  & \tt xvp          & 29.9   & 8.24   & 83.8\%     \\
           & \tt xvp\_new     & \underline{3.16}   & \underline{0.872}  & \underline{8.87\%}     \\
           & \tt PID          & 0      & 0      &    0\%     \\
           & Subtotal         & 33.0   & 9.12   & 92.7\%     \\
\hline
Coarse mesh& \tt rho\_c       & 0.296  & 0.0818 & 0.831\%    \\
           & \tt vfield       & 0.889  & 0.245  & 2.49\%     \\
           & \tt kern\_c      & 0.340  & 0.0939 & 0.954\%    \\
           & \tt force\_c     & (0.690)  & (0.190)  & (1.94\%)    \\
           & \tt rho\_c\_new  & 0.0140 & 0.00388  & 0.0394\% \\
           & Pencil-FFT       & (0.454)  & (0.125) & (1.27\%)      \\
           & Subtotal         & 1.55   & 0.429 & 4.36\%     \\
\hline
Fine mesh  & \tt kern\_f      & 1.06   & 0.292 & 2.97\%      \\
           & \tt force\_f     & (1.63) & (0.450) & (4.57\%)     \\
           & Fine-FFT     &   (1.41) & (0.389) & (3.96\%)     \\
           & Subtotal      & 1.06   & 0.292 & 2.97	\%      \\
\hline
Total &                      & 35.6 & 9.84 & 100\%\\
\hline
\multicolumn{2}{l} {Optimal limit} & 21.7 & 6 & 61.0\%\\
\hline
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
  \includegraphics[width=0.95\linewidth]{power_universe1-8_z0and1.pdf}
 \caption{We don't need this figure.}
\label{fig.power}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=1.1\linewidth]{diff_power.pdf}
 \caption{Fractional error on power spectrum and cross correlations.}
\label{fig.ccc}
\end{figure}
